---
title: "Flow Cytometry Analysis"
author: "Maggi Brisbin"
date: "7/25/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = TRUE, warning = FALSE, 
                      message = FALSE, cache.lazy = FALSE)
```

# packages

```{r}
library(flowCore)
library(ggcyto)
library(tidyverse)
library(scales)
library(growthrates)
library(ggpubr)
library(flowStats)
library(kableExtra)
library(magick)

"%ni%" <- Negate("%in%")
```


# Trial 1

Load Flow Cytometry data:
```{r load trial 1 data, cache=TRUE}
rnd1fs1<- read.flowSet(path = "20220803_BACTB12_3_1_1200/", pattern = ".fcs",alter.names = T)
rnd1fs2 <- read.flowSet(path = "20220803_BactB12_3_1_1500/", pattern = ".fcs",alter.names = T)
rnd1fs3 <- read.flowSet(path = "20220803_bactB12_3_1_2100/", pattern = ".fcs",alter.names = T)
rnd1fs4 <- read.flowSet(path = "20220804_bactb12_3_2_0900/", pattern = ".fcs",alter.names = T)
rnd1fs5 <- read.flowSet(path = "20220804_bactb12_3_2_1200/", pattern = ".fcs",alter.names = T)
```

Add sample names for each run:
```{r}
samples <- c( "Rpom_full_cold_1", "Rpom_full_cold_2", "Rpom_full_cold_3","Amac_full_cold_1", "Amac_full_cold_2", "Amac_full_cold_3",  "Rpom_noB12_cold_1", "Rpom_noB12_cold_2", "Rpom_noB12_cold_3", "Amac_noB12_cold_1", "Amac_noB12_cold_2","Amac_noB12_cold_3",
"Rpom_full_med_1", "Rpom_full_med_2", "Rpom_full_med_3", "Amac_full_med_1", "Amac_full_med_2", "Amac_full_med_3",
"Rpom_noB12_med_1", "Rpom_noB12_med_2", "Rpom_noB12_med_3", "Amac_noB12_med_1", "Amac_noB12_med_2", "Amac_noB12_med_3",
"Rpom_full_hot_1", "Rpom_full_hot_2", "Rpom_full_hot_3", "Amac_full_hot_1", "Amac_full_hot_2", "Amac_full_hot_3",
"Rpom_noB12_hot_1", "Rpom_noB12_hot_2", "Rpom_noB12_hot_3","Amac_noB12_hot_1", "Amac_noB12_hot_2", "Amac_noB12_hot_3")

pData(rnd1fs1)$well <- paste0(samples, "_1")
pData(rnd1fs2)$well <- paste0(samples, "_2")
pData(rnd1fs3)$well <- paste0(samples, "_3")
pData(rnd1fs4)$well <- paste0(samples, "_4")
pData(rnd1fs5)$well <- paste0(samples, "_5")

```

Create gate for live cells:
```{r}
g.singlets <- polygonGate(filterId = "Singlets","FL1.A"=c(3e3,1e6,1e6, 3e3),"FL3.A"=c(0,5000, 1e5, 1200)) # define gate
```

Create log version of gate for plotting and plot log transformed data with log transformed gate:
```{r}
logg.singlets <- polygonGate(filterId = "Singlets", "FL1.A"=c(log(3e3),log(1e6), log(1e6), log(3e3)),"FL3.A"=c(4, log(5000), log(1e5), log(1200))) 

autoplot(transform(rnd1fs1[[1]], `FL1.A` =log(`FL1.A` ), `FL3.A` =log(`FL3.A` ) ), "FL1.A","FL3.A", bins = 400, gate = logg.singlets) +geom_gate(logg.singlets) +geom_stats(type = c("percent", "count"))
```


Filter flowSet based on gate and then replot: (sanity check)
```{r}
filteredresult <- flowCore::Subset(rnd1fs1,g.singlets)

autoplot(transform(filteredresult[[1]], `FL1.A` =log(`FL1.A` ), `FL3.A` =log(`FL3.A` ) ), "FL1.A","FL3.A", bins = 400, gate = logg.singlets) +geom_gate(logg.singlets) +geom_stats(type = c("percent"))
```


```{r}
autoplot(transform(filteredresult, `FL1.A` =log(`FL1.A` ), `FL3.A` =log(`FL3.A` ) ), "FL1.A","FL3.A", bins = 400, gate = logg.singlets) +geom_gate(logg.singlets) 
```
Make a forward scatter gate:
```{r}
rg1 <- rectangleGate("FSC.A"=c(100, 100000), filterId="NonDebris")

```
Check forward scatter gate:
```{r}
autoplot(rnd1fs1[[5]], x = "FSC.A") +geom_gate(rg1) +geom_stats()
```

Filter based on forward scatter: 
```{r, cache=TRUE}
filteredresult <- flowCore::Subset(rnd1fs1,g.singlets)
filteredresult2 <- flowCore::Subset(filteredresult,rg1)
```

plot filtered forward scatter (sanity check):
```{r}
ggcyto(filteredresult2, aes(x = "FSC.A")) +geom_gate(rg1) +geom_density()
```

Get the total volume injected into the instrument for each sample at each time point and save as a character vector: 
```{r}
rnd1vols1<- c()
for (i in 1:length(samples)) {
  rnd1vols1[i] = rnd1fs1[[i]]@description["$VOL"][[1]]
}

rnd1vols2<- c()
for (i in 1:length(samples)) {
  rnd1vols2[i] = rnd1fs2[[i]]@description["$VOL"][[1]]
}

rnd1vols3<- c()
for (i in 1:length(samples)) {
  rnd1vols3[i] = rnd1fs3[[i]]@description["$VOL"][[1]]
}

rnd1vols4<- c()
for (i in 1:length(samples)) {
  rnd1vols4[i] = rnd1fs4[[i]]@description["$VOL"][[1]]
}

rnd1vols5<- c()
for (i in 1:length(samples)) {
  rnd1vols5[i] = rnd1fs5[[i]]@description["$VOL"][[1]]
}
```

Create gatingset from filtered flowset to get event counts for each treatment at each time point.
Add a volume column using the volume vectors for each day. 
```{r, cache=TRUE}
#1
filtered_gs1 <- GatingSet(filteredresult2)
gs_pop_add(filtered_gs1,g.singlets) # add gate to GatingSet
recompute(filtered_gs1)

filtered_ps1 <- gs_pop_get_count_with_meta(filtered_gs1)
filtered_ps1$vol <- rnd1vols1

#2
fr_2 <- flowCore::Subset(rnd1fs2,g.singlets)
fr2_2 <- flowCore::Subset(fr_2,rg1)

filtered_gs2 <- GatingSet(fr2_2)
gs_pop_add(filtered_gs2,g.singlets) # add gate to GatingSet
recompute(filtered_gs2)

filtered_ps2 <- gs_pop_get_count_with_meta(filtered_gs2)
filtered_ps2$vol <- rnd1vols2

#3
fr_3 <- flowCore::Subset(rnd1fs3,g.singlets)
fr2_3 <- flowCore::Subset(fr_3,rg1)

filtered_gs3 <- GatingSet(fr2_3)
gs_pop_add(filtered_gs3, g.singlets) # add gate to GatingSet
recompute(filtered_gs3)

filtered_ps3 <- gs_pop_get_count_with_meta(filtered_gs3)
filtered_ps3$vol <- rnd1vols3

#4
fr_4 <- flowCore::Subset(rnd1fs4,g.singlets)
fr2_4 <- flowCore::Subset(fr_4,rg1)

filtered_gs4 <- GatingSet(fr2_4)
gs_pop_add(filtered_gs4,g.singlets) # add gate to GatingSet
recompute(filtered_gs4)

filtered_ps4 <- gs_pop_get_count_with_meta(filtered_gs4)
filtered_ps4$vol <- rnd1vols4

#5

fr_5 <- flowCore::Subset(rnd1fs5,g.singlets)
fr2_5 <- flowCore::Subset(fr_5,rg1)

filtered_gs5 <- GatingSet(fr2_5)
gs_pop_add(filtered_gs5,g.singlets) # add gate to GatingSet
recompute(filtered_gs5)

filtered_ps5 <- gs_pop_get_count_with_meta(filtered_gs5)
filtered_ps5$vol <- rnd1vols5
```

Bind together count frames and create new column for volume in µl. Divide counts by volume to normalize. Separate names to get columns for grouping and faceting. Add factors for temperature. 
```{r}
results_t1 <- rbind(filtered_ps1, filtered_ps2, filtered_ps3, filtered_ps4, filtered_ps5)

results_t1$vol_ul <- as.numeric(results_t1$vol) / 1000

results_t1$cells_per_ul  <-   results_t1$Count /  results_t1$vol_ul


results_t1 <- results_t1 %>% separate(well,c("Species", "Media", "Temp", "Rep", "Time_point") )
results_t1$Temp <- factor(results_t1$Temp, levels = c("cold", "med", "hot"))
```

### Growth Curve Plot
Convert "time points" to hours and plot growth curve: 
```{r}
r2_t1<- results_t1 %>% 
  mutate(hours = case_when(
    Time_point == 1 ~ 3,
    Time_point == 2 ~ 6, 
    Time_point == 3 ~ 12,
    Time_point == 4 ~ 24,
    Time_point == 5 ~ 27)) 

r2_t1 %>% 
  group_by(hours, Media, Species, Temp) %>%
  dplyr::summarize(mean_cells_per_ul = mean(cells_per_ul), sd = sd(cells_per_ul)) %>% ggplot(aes(x = hours, y = mean_cells_per_ul, color = Temp, shape = Media)) + facet_grid(Temp~Species) + geom_point(size =2, alpha = .7) + theme_bw() + scale_color_manual(values = c("#2D55CC", "#FAAB33", "#FF4041"))  +geom_line(aes(linetype = Media))+
  geom_errorbar(aes(ymin=mean_cells_per_ul-sd, ymax=mean_cells_per_ul+sd), width=0) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
```
ggsave("results2_means.pdf", width = 4, height = 5)

#### log scale

plot with Log scale on y-axis:
```{r}
r2_t1 %>% 
  group_by(hours, Media, Species, Temp) %>%
  dplyr::summarize(mean_cells_per_ul = mean(cells_per_ul), sd = sd(cells_per_ul)) %>% ggplot(aes(x = hours, y = mean_cells_per_ul, color = Temp, shape = Media)) + facet_grid(Temp~Species) + geom_point(size =2, alpha = .7) + theme_bw() + scale_color_manual(values = c("#2D55CC", "#FAAB33", "#FF4041")) +
    geom_line(aes(linetype = Media))+
  geom_errorbar(aes(ymin=mean_cells_per_ul-sd, ymax=mean_cells_per_ul+sd), width=0) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+ scale_y_log10()
```
ggsave("t1_log_growthcurve.pdf", width = 7, height = 6)

## Maximum cell counts
Get and plot maximum cell concentrations:

```{r}
df_mean <- data.frame(r2_t1 %>% 
  group_by(Media, Species, Temp, Rep) %>%
  dplyr::summarize(max_cells_per_ul = max(cells_per_ul)) %>% 
  dplyr::summarize(mean_max_cells = mean(max_cells_per_ul), sd = sd(max_cells_per_ul)) %>% ungroup())

df_sd <-  data.frame(r2_t1 %>% 
  group_by(Media, Species, Temp, Rep) %>%
  dplyr::summarize(max_cells_per_ul = max(cells_per_ul)) %>% 
  dplyr::summarize(mean_max_cells = mean(max_cells_per_ul), sd = sd(max_cells_per_ul)) %>% ungroup() %>% mutate(ymin = mean_max_cells-sd) %>% mutate(ymax = mean_max_cells+sd))

df_points <- data.frame(r2_t1 %>% 
  group_by(Media, Species, Temp, Rep) %>%
  dplyr::summarize(max_cells_per_ul = max(cells_per_ul)) %>% ungroup())
  

ggplot() + geom_point(data = df_mean, aes(x = Media, y = mean_max_cells, color = Media), size =10, shape ="-") + theme_bw() + scale_color_manual(values = c( "black", "#6495ed"))+ geom_errorbar(data = df_sd, aes(x = Media, y = mean_max_cells, ymin=ymin, ymax=ymax, color = Media), width=0) + geom_jitter(data =df_points, aes(x =Media, y = max_cells_per_ul, color = Media), shape = 1, width = 0.114, fill = "white") +facet_grid(Species~ Temp, scales = "free_y") + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + ggtitle("Trial 1")

```
ggsave("t1_maxcells_mean_dots.pdf", width = 5, height = 6)


#### stats

```{r}
amac_biomass <- df_points %>% dplyr::filter(Species == "Amac")
amac_biomass$TempMedia <- paste(amac_biomass$Media, amac_biomass$Temp)
```

Amac cold
```{r}
amac_biomass_cold_noB12 <- amac_biomass %>% dplyr::filter(TempMedia == "noB12 cold")
amac_biomass_cold_full <- amac_biomass %>% dplyr::filter(TempMedia == "full cold")
t.test(amac_biomass_cold_noB12$max_cells_per_ul, amac_biomass_cold_full$max_cells_per_ul)
```
Amac Mid
```{r}
amac_biomass_mid_noB12 <- amac_biomass %>% dplyr::filter(TempMedia == "noB12 med")
amac_biomass_mid_full <- amac_biomass %>% dplyr::filter(TempMedia == "full med")
t.test(amac_biomass_mid_noB12$max_cells_per_ul, amac_biomass_mid_full$max_cells_per_ul)
```
Amac Hot
```{r}
amac_biomass_hot_noB12 <- amac_biomass %>% dplyr::filter(TempMedia == "noB12 hot")
amac_biomass_hot_full <- amac_biomass %>% dplyr::filter(TempMedia == "full hot")
t.test(amac_biomass_hot_noB12$max_cells_per_ul, amac_biomass_hot_full$max_cells_per_ul)
```

Rpom
```{r}
rpom_biomass <- df_points %>% dplyr::filter(Species == "Rpom")
rpom_biomass$TempMedia <- paste(rpom_biomass$Media, rpom_biomass$Temp)
```

Rpom cold
```{r}
rpom_biomass_cold_noB12 <- rpom_biomass %>% dplyr::filter(TempMedia == "noB12 cold")
rpom_biomass_cold_full <- rpom_biomass %>% dplyr::filter(TempMedia == "full cold")
t.test(rpom_biomass_cold_noB12$max_cells_per_ul, rpom_biomass_cold_full$max_cells_per_ul)
```
Rpom Mid
```{r}
rpom_biomass_mid_noB12 <- rpom_biomass %>% dplyr::filter(TempMedia == "noB12 med")
rpom_biomass_mid_full <- rpom_biomass %>% dplyr::filter(TempMedia == "full med")
t.test(rpom_biomass_mid_noB12$max_cells_per_ul, rpom_biomass_mid_full$max_cells_per_ul)
```

Rpom Hot
```{r}
rpom_biomass_hot_noB12 <- rpom_biomass %>% dplyr::filter(TempMedia == "noB12 hot")
rpom_biomass_hot_full <- rpom_biomass %>% dplyr::filter(TempMedia == "full hot")
t.test(rpom_biomass_hot_noB12$max_cells_per_ul, rpom_biomass_hot_full$max_cells_per_ul)
```


## cell size

#### mean

```{r, cache=TRUE}
FSC1<- data.frame(fsApply(filteredresult2,each_col,mean))
FSC2<- data.frame(fsApply(fr2_2,each_col,mean))
FSC3<- data.frame(fsApply(fr2_3,each_col,mean))
FSC4<- data.frame(fsApply(fr2_4,each_col,mean))
FSC5<- data.frame(fsApply(fr2_5,each_col,mean))

FSC1$well <- paste0(samples, "_1")
FSC2$well <- paste0(samples, "_2")
FSC3$well <- paste0(samples, "_3")
FSC4$well <- paste0(samples, "_4")
FSC5$well <- paste0(samples, "_5")

Fscatter_t1<- rbind(FSC1,FSC2,FSC3,FSC4,FSC5) %>% separate(well,c("Species", "Media", "Temp", "Rep", "Time_point"))

Fscatter_t1$Temp <- factor(Fscatter_t1$Temp, levels = c("cold", "med", "hot"))
```


```{r}
fscatt1<- Fscatter_t1 %>% 
  mutate(hours = case_when(
    Time_point == 1 ~ 3,
    Time_point == 2 ~ 6, 
    Time_point == 3 ~ 12,
    Time_point == 4 ~ 24,
    Time_point == 5 ~ 27)) 

fscatt1  %>% 
  group_by(hours, Media, Species, Temp) %>%
  dplyr::summarize(mean_size = mean(FSC.A), sd = sd(FSC.A)) %>% ggplot(aes(x = hours, y = mean_size, color = Temp, shape = Media)) + facet_grid(Temp~Species) + geom_point(size =2, alpha = .7) + theme_bw() + scale_color_manual(values = c("#2D55CC", "#FAAB33", "#FF4041"))  +geom_line(aes(linetype = Media))+
  geom_errorbar(aes(ymin=mean_size-sd, ymax=mean_size+sd), width=0) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
```
ggsave("t1_mean_FSC.pdf", width = 7, height = 6)

# Trial 2

Data import:
```{r, cache=TRUE}
rnd2fs1<- read.flowSet(path = "20220808_bactb12_4_1_0900/", pattern = ".fcs",alter.names = T)
rnd2fs2 <- read.flowSet(path = "20220808_BACTB12_4_1_1000//", pattern = ".fcs",alter.names = T)
rnd2fs3 <- read.flowSet(path = "20220808_bactb12_4_1_1100/", pattern = ".fcs",alter.names = T)
rnd2fs4 <- read.flowSet(path = "20220808_bactb12_4_1_1200/", pattern = ".fcs",alter.names = T)
rnd2fs5 <- read.flowSet(path = "20220808_bactb12_4_1_1300/", pattern = ".fcs",alter.names = T)
rnd2fs6 <- read.flowSet(path = "20220808_bactb12_4_1_1400/",  pattern = ".fcs",alter.names = T)
rnd2fs7 <- read.flowSet(path = "20220808_bactb12_4_1_1500/",  pattern = ".fcs",alter.names = T)
rnd2fs8 <- read.flowSet(path = "20220808_bactb12_4_1_1600/",  pattern = ".fcs",alter.names = T)
rnd2fs9 <- read.flowSet(path = "20220808_bactb12_4_1_1700/",  pattern = ".fcs",alter.names = T)
rnd2fs10 <- read.flowSet(path = "20220808_bactb12_4_1_2000/",  pattern = ".fcs",alter.names = T)
rnd2fs11 <- read.flowSet(path = "20220808_bactb12_4_1_2100/",  pattern = ".fcs",alter.names = T)
rnd2fs12 <- read.flowSet(path = "20220809_bactb12_4_2_0900/",  pattern = ".fcs",alter.names = T)
rnd2fs13 <- read.flowSet(path = "20220809_bactb12_4_2_0900_2/",  pattern = ".fcs",alter.names = T)

```

Add sample names:
```{r}
sampleNames3 <- c( "Rpom_full_cold_1", "Rpom_full_cold_2", "Rpom_full_cold_3","Amac_full_cold_1", "Amac_full_cold_2", "Amac_full_cold_3",  "Rpom_noB12_cold_1", "Rpom_noB12_cold_2", "Rpom_noB12_cold_3", "Amac_noB12_cold_1", "Amac_noB12_cold_2","Amac_noB12_cold_3",
"Rpom_full_med_1", "Rpom_full_med_2", "Rpom_full_med_3", "Amac_full_med_1", "Amac_full_med_2", "Amac_full_med_3",
"Rpom_noB12_med_1", "Rpom_noB12_med_2", "Rpom_noB12_med_3", "Amac_noB12_med_1", "Amac_noB12_med_2", "Amac_noB12_med_3",
"Rpom_full_hot_1", "Rpom_full_hot_2", "Rpom_full_hot_3", "Amac_full_hot_1", "Amac_full_hot_2", "Amac_full_hot_3",
"Rpom_noB12_hot_1", "Rpom_noB12_hot_2", "Rpom_noB12_hot_3","Amac_noB12_hot_1", "Amac_noB12_hot_2", "Amac_noB12_hot_3")

sampleNames4 <- c( "Rpom_full_cold_1", "Rpom_full_cold_2", "Rpom_full_cold_3","Amac_full_cold_1", "Amac_full_cold_2", "Amac_full_cold_3",  "Rpom_noB12_cold_1", "Rpom_noB12_cold_2", "Rpom_noB12_cold_3", "Amac_noB12_cold_1", "Amac_noB12_cold_2","Amac_noB12_cold_3",
"Amac_full_med_1", "Amac_full_med_2", "Amac_full_med_3",
"Amac_noB12_med_1", "Amac_noB12_med_2", "Amac_noB12_med_3",
"Amac_full_hot_1", "Amac_full_hot_2", "Amac_full_hot_3",
"Amac_noB12_hot_1", "Amac_noB12_hot_2", "Amac_noB12_hot_3")

sampleNames5 <- c( "Rpom_full_cold_1", "Rpom_full_cold_2", "Rpom_full_cold_3","Amac_full_cold_1", "Amac_full_cold_2", "Amac_full_cold_3",  "Rpom_noB12_cold_1", "Rpom_noB12_cold_2", "Rpom_noB12_cold_3", "Amac_noB12_cold_1", "Amac_noB12_cold_2","Amac_noB12_cold_3")

sampleNames6 <- c( "Amac_full_med_1", "Amac_full_med_2", "Amac_full_med_3",
"Amac_noB12_med_1", "Amac_noB12_med_2", "Amac_noB12_med_3",
"Amac_full_hot_1", "Amac_full_hot_2", "Amac_full_hot_3",
"Amac_noB12_hot_1", "Amac_noB12_hot_2", "Amac_noB12_hot_3")

pData(rnd2fs1)$well <- paste0(sampleNames3, "_1")
pData(rnd2fs2)$well <- paste0(sampleNames3, "_2")
pData(rnd2fs3)$well <- paste0(sampleNames3, "_3")
pData(rnd2fs4)$well <- paste0(sampleNames3, "_4")
pData(rnd2fs5)$well <- paste0(sampleNames3, "_5")
pData(rnd2fs6)$well <- paste0(sampleNames3, "_6")
pData(rnd2fs7)$well <- paste0(sampleNames3, "_7")
pData(rnd2fs8)$well <- paste0(sampleNames3, "_8")
pData(rnd2fs9)$well <- paste0(sampleNames3, "_9")
pData(rnd2fs10)$well <- paste0(sampleNames4, "_10")
pData(rnd2fs11)$well <- paste0(sampleNames4, "_11")
pData(rnd2fs12)$well <- paste0(sampleNames5, "_12")
pData(rnd2fs13)$well <- paste0(sampleNames6, "_13")

```

Double check gates:
```{r}
autoplot(transform(rnd2fs1[[1]], `FL1.A` =log(`FL1.A` ), `FL3.A` =log(`FL3.A` ) ), "FL1.A","FL3.A", bins = 400, gate = logg.singlets) +geom_gate(logg.singlets) +geom_stats(type = c("percent", "count"))
```


Filter flowSet based on gate and then replot: (sanity check)
```{r}
filteredresult <- flowCore::Subset(rnd2fs1, g.singlets)

autoplot(transform(filteredresult[[1]], `FL1.A` =log(`FL1.A` ), `FL3.A` =log(`FL3.A` ) ), "FL1.A","FL3.A", bins = 400, gate = logg.singlets) +geom_gate(logg.singlets) +geom_stats(type = c("percent"))
```

Forward scatter gate:
```{r}
autoplot(rnd2fs1[[1]], x = "FSC.A") +geom_gate(rg1) +geom_stats()
```

Filter flowSet based on gate and then replot: (sanity check)
```{r}
filteredresult <- flowCore::Subset(rnd2fs1,g.singlets)
filteredresult_2 <- flowCore::Subset(filteredresult, rg1)
```

```{r}
autoplot(filteredresult_2[[1]], x = "FSC.A") +geom_gate(rg1)
```

```{r}
ggcyto(filteredresult_2, aes(x = "FSC.A")) +geom_gate(rg1) +geom_density()
```
Get sample volumes for each time point:
```{r sample volumes}
rnd2vols1<- c()
for (i in 1:length(sampleNames3)) {
  rnd2vols1[i] = rnd2fs1[[i]]@description["$VOL"][[1]]
}

rnd2vols2<- c()
for (i in 1:length(sampleNames3)) {
  rnd2vols2[i] = rnd2fs2[[i]]@description["$VOL"][[1]]
}

rnd2vols3<- c()
for (i in 1:length(sampleNames3)) {
  rnd2vols3[i] = rnd2fs3[[i]]@description["$VOL"][[1]]
}

rnd2vols4<- c()
for (i in 1:length(sampleNames3)) {
  rnd2vols4[i] = rnd2fs4[[i]]@description["$VOL"][[1]]
}

rnd2vols5<- c()
for (i in 1:length(sampleNames3)) {
  rnd2vols5[i] = rnd2fs5[[i]]@description["$VOL"][[1]]
}

rnd2vols6<- c()
for (i in 1:length(sampleNames3)) {
  rnd2vols6[i] = rnd2fs6[[i]]@description["$VOL"][[1]]
}

rnd2vols7<- c()
for (i in 1:length(sampleNames3)) {
  rnd2vols7[i] = rnd2fs7[[i]]@description["$VOL"][[1]]
}

rnd2vols8<- c()
for (i in 1:length(sampleNames3)) {
  rnd2vols8[i] = rnd2fs8[[i]]@description["$VOL"][[1]]
}

rnd2vols9<- c()
for (i in 1:length(sampleNames3)) {
  rnd2vols9[i] = rnd2fs9[[i]]@description["$VOL"][[1]]
}

rnd2vols10<- c()
for (i in 1:length(sampleNames4)) {
  rnd2vols10[i] = rnd2fs10[[i]]@description["$VOL"][[1]]
}

rnd2vols11<- c()
for (i in 1:length(sampleNames4)) {
  rnd2vols11[i] = rnd2fs11[[i]]@description["$VOL"][[1]]
}

rnd2vols12<- c()
for (i in 1:length(sampleNames5)) {
  rnd2vols12[i] = rnd2fs12[[i]]@description["$VOL"][[1]]
}

rnd2vols13<- c()
for (i in 1:length(sampleNames6)) {
  rnd2vols13[i] = rnd2fs13[[i]]@description["$VOL"][[1]]
}
```


create gating sets for each timepoint:
```{r, cache=TRUE}
filtered_gs1 <- GatingSet(filteredresult_2)
gs_pop_add(filtered_gs1,g.singlets) # add gate to GatingSet
recompute(filtered_gs1)

filtered_ps1 <- gs_pop_get_count_with_meta(filtered_gs1)
filtered_ps1$vol <- rnd2vols1

fr_2 <- flowCore::Subset(rnd2fs2,g.singlets)
fr2_2 <- flowCore::Subset(fr_2,rg1)

filtered_gs2 <- GatingSet(fr2_2)
gs_pop_add(filtered_gs2,g.singlets) # add gate to GatingSet
recompute(filtered_gs2)

filtered_ps2 <- gs_pop_get_count_with_meta(filtered_gs2)
filtered_ps2$vol <- rnd2vols2

#3
fr_3 <- flowCore::Subset(rnd2fs3,g.singlets)
fr2_3 <- flowCore::Subset(fr_3,rg1)

filtered_gs3 <- GatingSet(fr2_3)
gs_pop_add(filtered_gs3, g.singlets) # add gate to GatingSet
recompute(filtered_gs3)

filtered_ps3 <- gs_pop_get_count_with_meta(filtered_gs3)
filtered_ps3$vol <- rnd2vols3

#4
fr_4 <- flowCore::Subset(rnd2fs4,g.singlets)
fr2_4 <- flowCore::Subset(fr_4,rg1)

filtered_gs4 <- GatingSet(fr2_4)
gs_pop_add(filtered_gs4,g.singlets) # add gate to GatingSet
recompute(filtered_gs4)

filtered_ps4 <- gs_pop_get_count_with_meta(filtered_gs4)
filtered_ps4$vol <- rnd2vols4

#5

fr_5 <- flowCore::Subset(rnd2fs5,g.singlets)
fr2_5 <- flowCore::Subset(fr_5,rg1)

filtered_gs5 <- GatingSet(fr2_5)
gs_pop_add(filtered_gs5,g.singlets) # add gate to GatingSet
recompute(filtered_gs5)

filtered_ps5 <- gs_pop_get_count_with_meta(filtered_gs5)
filtered_ps5$vol <- rnd2vols5

#6

fr_6 <- flowCore::Subset(rnd2fs6,g.singlets)
fr2_6 <- flowCore::Subset(fr_6,rg1)

filtered_gs6 <- GatingSet(fr2_6)
gs_pop_add(filtered_gs6,g.singlets) # add gate to GatingSet
recompute(filtered_gs6)

filtered_ps6 <- gs_pop_get_count_with_meta(filtered_gs6)
filtered_ps6$vol <- rnd2vols6

#7

fr_7 <- flowCore::Subset(rnd2fs7,g.singlets)
fr2_7 <- flowCore::Subset(fr_7,rg1)

filtered_gs7 <- GatingSet(fr2_7)
gs_pop_add(filtered_gs7,g.singlets) # add gate to GatingSet
recompute(filtered_gs7)

filtered_ps7 <- gs_pop_get_count_with_meta(filtered_gs7)
filtered_ps7$vol <- rnd2vols7

#8

fr_8 <- flowCore::Subset(rnd2fs8,g.singlets)
fr2_8 <- flowCore::Subset(fr_8,rg1)

filtered_gs8  <- GatingSet(fr2_8)
gs_pop_add(filtered_gs8,g.singlets) # add gate to GatingSet
recompute(filtered_gs8)

filtered_ps8 <- gs_pop_get_count_with_meta(filtered_gs8)
filtered_ps8$vol <- rnd2vols8

#9

fr_9 <- flowCore::Subset(rnd2fs9,g.singlets)
fr2_9 <- flowCore::Subset(fr_9,rg1)

filtered_gs9 <- GatingSet(fr2_9)
gs_pop_add(filtered_gs9,g.singlets) # add gate to GatingSet
recompute(filtered_gs9)

filtered_ps9 <- gs_pop_get_count_with_meta(filtered_gs9)
filtered_ps9$vol <- rnd2vols9

#10

fr_10 <- flowCore::Subset(rnd2fs10,g.singlets)
fr2_10 <- flowCore::Subset(fr_10,rg1)

filtered_gs10 <- GatingSet(fr2_10)
gs_pop_add(filtered_gs10,g.singlets) # add gate to GatingSet
recompute(filtered_gs10)

filtered_ps10 <- gs_pop_get_count_with_meta(filtered_gs10)
filtered_ps10$vol <- rnd2vols10

#11

fr_11 <- flowCore::Subset(rnd2fs11,g.singlets)
fr2_11 <- flowCore::Subset(fr_11,rg1)

filtered_gs11 <- GatingSet(fr2_11)
gs_pop_add(filtered_gs11,g.singlets) # add gate to GatingSet
recompute(filtered_gs11)

filtered_ps11 <- gs_pop_get_count_with_meta(filtered_gs11)
filtered_ps11$vol <- rnd2vols11

#12

fr_12 <- flowCore::Subset(rnd2fs12,g.singlets)
fr2_12 <- flowCore::Subset(fr_12,rg1)

filtered_gs12 <- GatingSet(fr2_12)
gs_pop_add(filtered_gs12,g.singlets) # add gate to GatingSet
recompute(filtered_gs12)

filtered_ps12 <- gs_pop_get_count_with_meta(filtered_gs12)
filtered_ps12$vol <- rnd2vols12

#13

fr_13 <- flowCore::Subset(rnd2fs13,g.singlets)
fr2_13 <- flowCore::Subset(fr_13,rg1)

filtered_gs13 <- GatingSet(fr2_13)
gs_pop_add(filtered_gs13,g.singlets) # add gate to GatingSet
recompute(filtered_gs13)

filtered_ps13 <- gs_pop_get_count_with_meta(filtered_gs13)
filtered_ps13$vol <- rnd2vols13
```
## Growth Curve Plot 
Put together time points into single dataframe, compute concentrations, split sample names for grouping and faceting: 
```{r}
results <- rbind(filtered_ps1, filtered_ps2, filtered_ps3, filtered_ps4, filtered_ps5, filtered_ps6, filtered_ps7, filtered_ps8, filtered_ps9, filtered_ps10, filtered_ps11, filtered_ps12, filtered_ps13)

results$vol_ul <- as.numeric(results$vol) / 1000

results$cells_per_ul  <-   results$Count /  results$vol_ul


results <- results %>% separate(well,c("Species", "Media", "Temp", "Rep", "Time_point") )
results$Temp <- factor(results$Temp, levels = c("cold", "med", "hot"))
```

convert time points to hours and plot!
```{r}
r2<- results %>% 
  mutate(hours = case_when(
    Time_point == 1 ~ 0,
    Time_point == 2 ~ 1, 
    Time_point == 3 ~ 2,
    Time_point == 4 ~ 3,
    Time_point == 5 ~ 4,
    Time_point == 6 ~ 5,
    Time_point == 7 ~ 6,
    Time_point == 8 ~ 7,
    Time_point == 9 ~ 8,
    Time_point == 10 ~ 9,
    Time_point == 11 ~ 12,
    Time_point == 12 ~ 24,
    Time_point == 13 ~ 24
    ))

r2 %>% group_by(hours, Media, Species, Temp) %>%
  dplyr::summarize(mean_cells_per_ul = mean(cells_per_ul), sd = sd(cells_per_ul)) %>% ggplot(aes(x = hours, y = mean_cells_per_ul, color = Temp, shape = Media)) + facet_grid(Temp~Species) + geom_point(size =2, alpha = .7) + theme_bw() + scale_color_manual(values = c("#2D55CC", "#FAAB33", "#FF4041")) + geom_line(aes(linetype = Media))+
  geom_errorbar(aes(ymin=mean_cells_per_ul-sd, ymax=mean_cells_per_ul+sd), width=0) + scale_y_continuous(labels = function(x) format(x, scientific = TRUE))

```

# log scale

```{r}
r2 %>% 
  group_by(hours, Media, Species, Temp) %>% 
  dplyr::summarize(mean_cells_per_ul = mean(cells_per_ul), sd = sd(cells_per_ul)) %>% dplyr::filter(!(Temp == "cold" &hours ==2)) %>% dplyr::filter(!(Temp == "hot" & hours ==1)) %>% ggplot(aes(x = hours, y = mean_cells_per_ul, color = Temp, shape = Media)) + facet_grid(Temp~Species) + geom_point(size =2, alpha = .7) + theme_light() + scale_color_manual(values = c("#2D55CC", "#FAAB33", "#FF4041")) +
   geom_line(aes(linetype = Media))+
  geom_errorbar(aes(ymin=mean_cells_per_ul-sd, ymax=mean_cells_per_ul+sd), width=0) + scale_y_log10() + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + xlim(0, 26)
```
ggsave("t2_log_growthcurve.pdf", width = 7, height = 6)


```{r}
results_Rpom_Hot_Full <- r2 %>% group_by(hours, Media, Species, Temp) %>%
  dplyr::summarize(mean_cells_per_ul = mean(cells_per_ul), sd = sd(cells_per_ul)) %>% dplyr::filter(Media == "full") %>%  dplyr::filter(Species == "Amac") %>% dplyr::filter(Temp == "hot")

results_Rpom_Hot_noB12 <- r2 %>% group_by(hours, Media, Species, Temp) %>%
  dplyr::summarize(mean_cells_per_ul = mean(cells_per_ul), sd = sd(cells_per_ul)) %>% dplyr::filter(Media == "noB12") %>%  dplyr::filter(Species == "Amac") %>% dplyr::filter(Temp == "hot")
```

biomass reduction 

```{r}
(35815 - 27738) / 35815
```


## maximum cell counts

```{r}
df_mean <- data.frame(r2 %>% 
  group_by(Media, Species, Temp, Rep) %>%
  dplyr::summarize(max_cells_per_ul = max(cells_per_ul)) %>% 
  dplyr::summarize(mean_max_cells = mean(max_cells_per_ul), sd = sd(max_cells_per_ul)) %>% ungroup())

df_sd <-  data.frame(r2 %>% 
  group_by(Media, Species, Temp, Rep) %>%
  dplyr::summarize(max_cells_per_ul = max(cells_per_ul)) %>% 
  dplyr::summarize(mean_max_cells = mean(max_cells_per_ul), sd = sd(max_cells_per_ul)) %>% ungroup() %>% mutate(ymin = mean_max_cells-sd) %>% mutate(ymax = mean_max_cells+sd))

df_points <- data.frame(r2 %>% 
  group_by(Media, Species, Temp, Rep) %>%
  dplyr::summarize(max_cells_per_ul = max(cells_per_ul)) %>% ungroup())
  

 ggplot() + geom_point(data = df_mean, aes(x = Media, y = mean_max_cells, color = Media), size =10, shape ="-") + theme_bw() + scale_color_manual(values = c( "black", "#6495ed"))+ geom_errorbar(data = df_sd, aes(x = Media, y = mean_max_cells, ymin=ymin, ymax=ymax, color = Media), width=0) + geom_jitter(data =df_points, aes(x =Media, y = max_cells_per_ul, color = Media), shape = 1, width = 0.114, fill = "white") +facet_grid(Species~ Temp, scales = "free_y") + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + ggtitle("Trial 2")

```
ggsave("t2_maxcells_mean_dots.pdf", width = 5, height = 6)

### stats

```{r}
amac_biomass <- df_points %>% dplyr::filter(Species == "Amac")
amac_biomass$TempMedia <- paste(amac_biomass$Media, amac_biomass$Temp)
```

Amac cold
```{r}
amac_biomass_cold_noB12 <- amac_biomass %>% dplyr::filter(TempMedia == "noB12 cold")
amac_biomass_cold_full <- amac_biomass %>% dplyr::filter(TempMedia == "full cold")
t.test(amac_biomass_cold_noB12$max_cells_per_ul, amac_biomass_cold_full$max_cells_per_ul)
```
Amac Mid
```{r}
amac_biomass_mid_noB12 <- amac_biomass %>% dplyr::filter(TempMedia == "noB12 med")
amac_biomass_mid_full <- amac_biomass %>% dplyr::filter(TempMedia == "full med")
t.test(amac_biomass_mid_noB12$max_cells_per_ul, amac_biomass_mid_full$max_cells_per_ul)
```
Amac Hot
```{r}
amac_biomass_hot_noB12 <- amac_biomass %>% dplyr::filter(TempMedia == "noB12 hot")
amac_biomass_hot_full <- amac_biomass %>% dplyr::filter(TempMedia == "full hot")
t.test(amac_biomass_hot_noB12$max_cells_per_ul, amac_biomass_hot_full$max_cells_per_ul)
```

Rpom
```{r}
rpom_biomass <- df_points %>% dplyr::filter(Species == "Rpom")
rpom_biomass$TempMedia <- paste(rpom_biomass$Media, rpom_biomass$Temp)
```

Rpom cold
```{r}
rpom_biomass_cold_noB12 <- rpom_biomass %>% dplyr::filter(TempMedia == "noB12 cold")
rpom_biomass_cold_full <- rpom_biomass %>% dplyr::filter(TempMedia == "full cold")
t.test(rpom_biomass_cold_noB12$max_cells_per_ul, rpom_biomass_cold_full$max_cells_per_ul)
```
Rpom Mid
```{r}
rpom_biomass_mid_noB12 <- rpom_biomass %>% dplyr::filter(TempMedia == "noB12 med")
rpom_biomass_mid_full <- rpom_biomass %>% dplyr::filter(TempMedia == "full med")
t.test(rpom_biomass_mid_noB12$max_cells_per_ul, rpom_biomass_mid_full$max_cells_per_ul)
```

Rpom Hot
```{r}
rpom_biomass_hot_noB12 <- rpom_biomass %>% dplyr::filter(TempMedia == "noB12 hot")
rpom_biomass_hot_full <- rpom_biomass %>% dplyr::filter(TempMedia == "full hot")
t.test(rpom_biomass_hot_noB12$max_cells_per_ul, rpom_biomass_hot_full$max_cells_per_ul)
```


## maximum growth rates


All line fits: 
```{r}


all_line_fits <- all_easylinear(cells_per_ul ~ hours|Species+Temp+Rep+Media,
                                data = (r2 %>% dplyr::filter(!(Temp == "cold" & hours ==2)) %>%  dplyr::filter(!(Temp == "hot" & hours ==1))), h=3,  quota = 0.95)

par(mfrow = c(6, 6))
par(mar = c(2.5, 4, 2, 1))
plot(all_line_fits, log = "y")
```

Rpom: 

```{r}
rpom_line_fits <- all_easylinear(cells_per_ul ~ hours|Species+Temp+Rep+Media,
                                data = (r2 %>% dplyr::filter(!(Temp == "cold" & hours ==2)) %>%  dplyr::filter(!(Temp == "hot" & hours ==1)) %>%  dplyr::filter(Species == "Rpom")), h=3,  quota = 0.95)

#pdf("rpom_line_fits.pdf")
par(mfrow = c(6, 3))
par(mar = c(2.5, 4, 2, 1))
plot(rpom_line_fits, log = "y")
#dev.off()
```

```{r}
kable((data.frame(results(rpom_line_fits)) %>% select(Species, Temp, Rep, Media, mumax, r2) %>% arrange(Temp)), format = "html", digits = 3) %>%
  kable_styling("striped") #%>% save_kable("rpom_line_fits.png", density = 10000)
```

Amac:

```{r}
amac_line_fits <- all_easylinear(cells_per_ul ~ hours|Species+Temp+Rep+Media,
                                data = (r2 %>% dplyr::filter(!(Temp == "cold" & hours ==2)) %>%  dplyr::filter(!(Temp == "hot" & hours ==1)) %>%  dplyr::filter(Species == "Amac")), h=3,  quota = 0.95)


#pdf("amac_line_fits.pdf")
par(mfrow = c(6, 3))
par(mar = c(2.5, 4, 2, 1))
plot(amac_line_fits, log = "y")
#dev.off()
```

```{r}
kable((data.frame(results(amac_line_fits)) %>% select(Species, Temp, Rep, Media, mumax, r2) %>% arrange(Temp)), format = "html", digits = 3) %>%
  kable_styling("striped") #%>% save_kable("amac_line_fits.png", density = 10000)
```


```{r}
#many_mus <- rbind( results(midhot_line_fits), results(cold_line_fits))
many_mus <- results(all_line_fits)
many_mus$Temp <- factor(many_mus$Temp, levels = c("cold" ,"med", "hot"), labels = c("cold", "mid", "hot"))
many_mus$TempMedia <- paste(many_mus$Temp, many_mus$Media)

my_comparisons = list( c("cold full", "cold noB12"), c("mid full", "mid noB12"), c("hot full", "hot noB12"))

many_mus %>% dplyr::filter(Species == "Amac") %>% ggplot(aes(x=Temp, y = mumax, color= Media)) + facet_grid(~Species) +geom_boxplot() + theme_bw() + scale_color_manual(values = c( "black", "#6495ed")) + theme(legend.position = "bottom")+ stat_compare_means(label.y = c(0.8, 1.0, 1.6), method = "t.test", group.by = "Temp")

#label = "p.signif"
```


```{r}
df_mean <- data.frame(many_mus %>% dplyr::filter(Species == "Amac") %>% group_by(Species, Temp, Media) %>%
  dplyr::summarize(mean_mumax = mean(mumax), sd = sd(mumax)) %>% ungroup())

df_sd <-  data.frame(many_mus %>% dplyr::filter(Species == "Amac") %>% group_by(Species, Temp, Media) %>%
  dplyr::summarize(mean_mumax = mean(mumax), sd = sd(mumax)) %>% ungroup()) %>% mutate(ymin = mean_mumax-sd) %>% mutate(ymax = mean_mumax+sd)

df_points <- data.frame(many_mus %>% dplyr::filter(Species == "Amac"))
  

amac_mumax<- ggplot() + geom_point(data = df_mean, aes(x = Media, y = mean_mumax, color = Media), size =10, shape ="-") + theme_bw() + scale_color_manual(values = c( "black", "#6495ed"))+ geom_errorbar(data = df_sd, aes(x = Media, y = mean_mumax, ymin=ymin, ymax=ymax, color = Media), width=0) + geom_jitter(data =df_points, aes(x =Media, y = mumax, color = Media), shape = 1, width = 0.114, fill = "white") +facet_grid(~Temp) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + ggtitle("A. macleodii") +ylim(0.5, 1.7)

amac_mumax
```


```{r}
many_mus %>% dplyr::filter(Species == "Rpom") %>% ggplot(aes(x=Temp, y = mumax, color= Media)) + facet_grid(~Species) +geom_boxplot() + theme_bw() + scale_color_manual(values = c( "black", "#6495ed")) + theme(legend.position = "bottom")+ stat_compare_means(label.y = c(0.8, 1.0, 1.6), method = "t.test", group.by = "Temp") + ylim(0.5, 1.8)
```


```{r}
df_mean <- data.frame(many_mus %>% dplyr::filter(Species == "Rpom") %>% group_by(Species, Temp, Media) %>%
  dplyr::summarize(mean_mumax = mean(mumax), sd = sd(mumax)) %>% ungroup())

df_sd <-  data.frame(many_mus %>% dplyr::filter(Species == "Rpom") %>% group_by(Species, Temp, Media) %>%
  dplyr::summarize(mean_mumax = mean(mumax), sd = sd(mumax)) %>% ungroup()) %>% mutate(ymin = mean_mumax-sd) %>% mutate(ymax = mean_mumax+sd)

df_points <- data.frame(many_mus %>% dplyr::filter(Species == "Rpom"))
  

rpom_mumax <- ggplot() + geom_point(data = df_mean, aes(x = Media, y = mean_mumax, color = Media), size =10, shape ="-") + theme_bw() + scale_color_manual(values = c( "black", "#6495ed"))+ geom_errorbar(data = df_sd, aes(x = Media, y = mean_mumax, ymin=ymin, ymax=ymax, color = Media), width=0) + geom_jitter(data =df_points, aes(x =Media, y = mumax, color = Media), shape = 1, width = 0.114, fill = "white") +facet_grid(~Temp) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + ggtitle("R. pomeroyi") +ylim(0.4, 1.9)

rpom_mumax
```

```{r}
library(patchwork)
(amac_mumax / rpom_mumax) + plot_layout(guides = "collect")
```

ggsave("mumax_mean_dots.pdf", width = 5, height = 6)


```{r}
many_mus %>% ggplot(aes(x=Temp, y = mumax, color= Media)) + facet_grid(~Species) +geom_boxplot() + theme_bw() + scale_color_manual(values = c( "black", "#6495ed")) + theme(legend.position = "bottom")+ stat_compare_means(label.y = c(0.8, 1.0, 1.6), method = "t.test", group.by = "Temp",  label = "p.signif") + ylim(0.5, 1.9) + theme(legend.position = "right")
```


```{r}
amac_mu <- many_mus %>% dplyr::filter(Species == "Amac")
amac_mu$TempMedia <- paste(amac_mu$Media, amac_mu$Temp)
```


```{r}
#cold
amac_mu_cold_noB12 <- amac_mu %>% dplyr::filter(TempMedia == "noB12 cold")
amac_mu_cold_full <- amac_mu %>% dplyr::filter(TempMedia == "full cold")
t.test(amac_mu_cold_noB12$mumax, amac_mu_cold_full$mumax)
```
```{r}
#warm
amac_mu_warm_noB12 <- amac_mu %>% dplyr::filter(TempMedia == "noB12 mid")
amac_mu_warm_full <- amac_mu %>% dplyr::filter(TempMedia == "full mid")
t.test(amac_mu_warm_noB12$mumax, amac_mu_warm_full$mumax)
```


```{r}
#warm
amac_mu_hot_noB12 <- amac_mu %>% dplyr::filter(TempMedia == "noB12 hot")
amac_mu_hot_full <- amac_mu %>% dplyr::filter(TempMedia == "full hot")
t.test(amac_mu_hot_noB12$mumax, amac_mu_hot_full$mumax)
```
Rpom Hot
```{r}
rpom_mu <- many_mus %>% dplyr::filter(Species == "Rpom")
rpom_mu_hot_noB12 <- rpom_mu %>% dplyr::filter(TempMedia == "hot noB12")
rpom_mu_hot_full <- rpom_mu %>% dplyr::filter(TempMedia == "hot full")
t.test(rpom_mu_hot_noB12$mumax, rpom_mu_hot_full$mumax)
```
Rpom Mid
```{r}
rpom_mu_mid_noB12 <- rpom_mu %>% dplyr::filter(TempMedia == "mid noB12")
rpom_mu_mid_full <- rpom_mu %>% dplyr::filter(TempMedia == "mid full")
t.test(rpom_mu_mid_noB12$mumax, rpom_mu_mid_full$mumax)
```
Rpom Cool
```{r}
rpom_mu_cool_noB12 <- rpom_mu %>% dplyr::filter(TempMedia == "cold noB12")
rpom_mu_cool_full <- rpom_mu %>% dplyr::filter(TempMedia == "cold full")
t.test(rpom_mu_cool_noB12$mumax, rpom_mu_cool_full$mumax)
```
## cell size 

### Mean
```{r, cache=TRUE}
FSC1<- data.frame(fsApply(filteredresult_2,each_col,mean))
FSC2<- data.frame(fsApply(fr2_2,each_col,mean))
FSC3<- data.frame(fsApply(fr2_3,each_col,mean))
FSC4<- data.frame(fsApply(fr2_4,each_col,mean))
FSC5<- data.frame(fsApply(fr2_5,each_col,mean))
FSC6<- data.frame(fsApply(fr2_6,each_col,mean))
FSC7<- data.frame(fsApply(fr2_7,each_col,mean))
FSC8<- data.frame(fsApply(fr2_8,each_col,mean))
FSC9<- data.frame(fsApply(fr2_9,each_col,mean))
FSC10<- data.frame(fsApply(fr2_10,each_col,mean))
FSC11<- data.frame(fsApply(fr2_11,each_col,mean))
FSC12<- data.frame(fsApply(fr2_12,each_col,mean))
FSC13<- data.frame(fsApply(fr2_13,each_col,mean))



FSC1$well <- paste0(sampleNames3, "_1")
FSC2$well <- paste0(sampleNames3, "_2")
FSC3$well <- paste0(sampleNames3, "_3")
FSC4$well <- paste0(sampleNames3, "_4")
FSC5$well <- paste0(sampleNames3, "_5")
FSC6$well <- paste0(sampleNames3, "_6")
FSC7$well <- paste0(sampleNames3, "_7")
FSC8$well <- paste0(sampleNames3, "_8")
FSC9$well <- paste0(sampleNames3, "_9")
FSC10$well <- paste0(sampleNames4, "_10")
FSC11$well <- paste0(sampleNames4, "_11")
FSC12$well <- paste0(sampleNames5, "_12")
FSC13$well <- paste0(sampleNames6, "_13")


Fscatter<- rbind(FSC1,FSC2,FSC3,FSC4,FSC5, FSC6, FSC7, FSC8, FSC9, FSC10, FSC11, FSC12, FSC13) %>% separate(well,c("Species", "Media", "Temp", "Rep", "Time_point"))

Fscatter$Temp <- factor(Fscatter$Temp, levels = c("cold", "med", "hot"))

fscat2 <- Fscatter %>% 
  mutate(hours = case_when(
    Time_point == 1 ~ 0,
    Time_point == 2 ~ 1, 
    Time_point == 3 ~ 2,
    Time_point == 4 ~ 3,
    Time_point == 5 ~ 4,
    Time_point == 6 ~ 5,
    Time_point == 7 ~ 6,
    Time_point == 8 ~ 7,
    Time_point == 9 ~ 8,
    Time_point == 10 ~ 9,
    Time_point == 11 ~ 12,
    Time_point == 12 ~ 24,
    Time_point == 13 ~ 24
    ))  

fscat2 %>% ggplot(aes(x = hours, y = FSC.A, color = Temp, shape = Media)) + facet_grid(Temp~Species) + geom_point(size =2, alpha = .7) + theme_light() + scale_color_manual(values = c("#2D55CC", "#FAAB33", "#FF4041")) +
    theme(
      strip.text.x = element_text(
        size = 12, color = "white", face = "bold.italic"
        )) + geom_line(aes(linetype = Media))
```

```{r}
fscat2  %>% 
  group_by(hours, Media, Species, Temp) %>%
  dplyr::summarize(mean_size = mean(FSC.A), sd = sd(FSC.A)) %>% ggplot(aes(x = hours, y = mean_size, color = Temp, shape = Media)) + facet_grid(Temp~Species) + geom_point(size =2, alpha = .7) + theme_bw() + scale_color_manual(values = c("#2D55CC", "#FAAB33", "#FF4041"))  +geom_line(aes(linetype = Media))+
  geom_errorbar(aes(ymin=mean_size-sd, ymax=mean_size+sd), width=0) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
```
ggsave("t2_mean_FSC.pdf", width = 7, height = 6)

```{r}
fscatt1$trial <- "1"
fscat2$trial <- "2"

fscatt1_2 <- rbind(fscatt1, fscat2) %>% filter(hours == 24) %>% filter(Temp == "hot") %>% filter(Species == "Amac")

```


# B12 Measurements

## final cell counts

to get number of cells at time of sacrifice to which to normalize b12 mas spec measurements

```{r}
results <- filtered_ps9

results$vol_ml <- as.numeric(results$vol) /1000 /1000

results$cells_per_ml  <-   results$Count /  results$vol_ml

results <- results %>% separate(well,c("Species", "Media", "Temp", "Rep", "Time_point") )
results$Temp <- factor(results$Temp, levels = c("cold", "med", "hot"))
```

```{r}
r4<- results %>% 
  mutate(hours = case_when(
    Time_point == 10 ~ 9)) %>% filter(Species == "Rpom") %>% filter(Media == "noB12") %>% filter(Temp != "cold")

r4$trial <- "2"

r4$Tube <- c(19,20,21,31,32,33)
```

```{r}
results <- filtered_ps12

results$vol_ml <- as.numeric(results$vol) /1000 /1000 #B12 data is mg/ml so getting cells /ml instedd of cells/ml

results$cells_per_ml  <-   results$Count /  results$vol_ml

results <- results %>% separate(well,c("Species", "Media", "Temp", "Rep", "Time_point") )
results$Temp <- factor(results$Temp, levels = c("cold", "med", "hot"))
```

```{r}
r5<- results %>% 
  mutate(hours = case_when(
    Time_point == 13 ~ 24)) %>% filter(Species == "Rpom") %>% filter(Media == "noB12") %>% filter(Temp == "cold")

r5$trial <- "2"

r5$Tube <- c(7,8,9)

df1 <- rbind(data.frame(r4 %>% group_by(hours, Media, Species, Temp) %>%
  dplyr::summarize(mean_cells_per_ml = mean(cells_per_ml), sd = sd(cells_per_ml)) %>% ungroup()), data.frame(r5 %>% group_by(hours, Media, Species, Temp) %>%
  dplyr::summarize(mean_cells_per_ml = mean(cells_per_ml), sd = sd(cells_per_ml)) %>% ungroup()))

df2 <- rbind(data.frame(r4 %>% group_by(hours, Media, Species, Temp) %>%
  dplyr::summarize(mean_cells_per_ml = mean(cells_per_ml), sd = sd(cells_per_ml)) %>% ungroup() %>% mutate(ymin = mean_cells_per_ml-sd) %>% mutate(ymax = mean_cells_per_ml+sd)), data.frame(r5 %>% group_by(hours, Media, Species, Temp) %>%
  dplyr::summarize(mean_cells_per_ml = mean(cells_per_ml), sd = sd(cells_per_ml)) %>% ungroup() %>% mutate(ymin = mean_cells_per_ml-sd) %>% mutate(ymax = mean_cells_per_ml+sd)))

df3 <- rbind(r4, r5)

ggplot() + geom_point(data = df1, aes(x = Temp, y = mean_cells_per_ml, color = Temp), size =10, shape ="-") + theme_light() + scale_color_manual(values = c("#2D55CC", "#FAAB33", "#FF4041")) +
    theme(
      strip.text.x = element_text(
        size = 12, color = "white", face = "bold.italic"
        )) +
  geom_errorbar(data = df2, aes(x = Temp, y = mean_cells_per_ml, ymin=ymin, ymax=ymax, color = Temp), width=0) + geom_jitter(data =df3, aes(x = Temp, y = cells_per_ml, color = Temp), shape = 1, width = 0.115)
```

```{r}
print(df3 %>% select(Species, Media, Temp, Rep, Tube, cells_per_ml))
```
These cell count values were added to the spreadsheet with B12 results.

## plot B12 concentrations normalized to cell counts 

```{r}
cob<- read.csv("cobalamin_supernatant_pellets_t2.csv")
str(cob)
```

normalize to mls and # of cells / ml

```{r}
cob$conc_mg_ml_in_extract <- cob$conc_mgL_in_final_extract / 1000
cob$mg_in_sample <- cob$conc_mg_ml_in_extract *  cob$Volume_mL
cob$cells_per_sample <- cob$Volume_mL * cob$cells.perml
cob$ug_per_cell <- cob$mg_in_sample / cob$cells_per_sample * 1000
```


```{r}

cob$Temp <- factor(cob$Temp, levels = c("cool", "warm", "hot"))

cob %>%  ggplot(aes(x=S_P, y=ug_per_cell)) + geom_boxplot(aes(color = Temp)) + scale_color_manual(values = c("#2D55CC", "#FAAB33", "#FF4041")) + theme_light() +
    theme(
      strip.text.x = element_text(
        size = 12, color = "white", face = "bold.italic"
        )) +
  stat_compare_means(label.y = c(3e-06), method = "t.test", group.by = "S_P")

```
ggsave("b12_normalized.pdf", width = 4, height = 3)

** convenient because of ggplot separates treatments, but boxplots are not appropriate for n=3 ... **

```{r}
df4 <- data.frame(cob %>% group_by(S_P, Temp) %>%
  dplyr::summarize(mean_ug_per_cell = mean(ug_per_cell), sd = sd(ug_per_cell)) %>% ungroup())

df5 <- data.frame(cob %>% group_by(S_P, Temp) %>%
  dplyr::summarize(mean_ug_per_cell = mean(ug_per_cell), sd = sd(ug_per_cell)) %>% ungroup()) %>% mutate(ymin = mean_ug_per_cell-sd) %>% mutate(ymax = mean_ug_per_cell+sd)

ggplot() + geom_point(data = df4, aes(x = Temp, y = mean_ug_per_cell, color = Temp), size =10, shape ="-") + theme_bw() + scale_color_manual(values = c("#2D55CC", "#FAAB33", "#FF4041")) + geom_errorbar(data = df5, aes(x = Temp, y = mean_ug_per_cell, ymin=ymin, ymax=ymax, color = Temp), width=0) + geom_jitter(data =cob, aes(x =Temp, y = ug_per_cell, color = Temp), shape = 1, width = 0.114, fill = "white") +facet_grid(~S_P) +ylim(	-1.124685e-07, 3e-06) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())

```

```{r}
ggplot() + geom_point(data = df4, aes(x = Temp, y = mean_ug_per_cell, color = Temp), size =12, shape ="-") + theme_light() + scale_color_manual(values = c("#2D55CC", "#FAAB33", "#FF4041")) + geom_errorbar(data = df5, aes(x = Temp, y = mean_ug_per_cell, ymin=ymin, ymax=ymax, color = Temp), width=0) + geom_dotplot(data =cob, aes(x =Temp, y = ug_per_cell, color = Temp), shape = 1, dotsize=0.5, binaxis='y', stackdir = 'center', fill = "white" ) +facet_grid(~S_P) +ylim(	-1.124685e-07, 3e-06) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
```





## Stats

Two-Way ANOVA - Temperature and Supernatant/Pellet

```{r}
cob$TempPellet <- paste(cob$Temp, cob$S_P)
anova_1 <- aov(ug_per_cell~Temp+S_P, data = cob)
summary(anova_1)
```
```{r}
TukeyHSD(anova_1, conf.level=.95)
```


```{r}
cob_coolP <- cob %>% dplyr::filter(Temp == "cool") %>% dplyr::filter(S_P == "P")
cob_warmP <- cob%>% dplyr::filter(Temp == "warm") %>% dplyr::filter(S_P == "P")

t.test(cob_coolP$ug_per_cell,cob_warmP$ug_per_cell)
```

```{r}
cob_hotP <- cob %>% dplyr::filter(Temp == "hot") %>% dplyr::filter(S_P == "P")

t.test(cob_coolP$ug_per_cell, cob_hotP$ug_per_cell)
```
```{r}
cob_S <- cob%>% dplyr::filter(S_P == "S")
cob_P <- cob %>% dplyr::filter(S_P == "P")

t.test(cob_P$ug_per_cell, cob_S$ug_per_cell)
```

